Asymmetric relationships and their effects on coexistence: Supporting information

### 2023-08-01 paper/RefereeResponses1.tex

\noindent C25) L95-99: I think this is a great teaser to get the readers to later dive in into your mathematics \\

\noindent ***\emph{Thank you! This is exactly our purpose, here, and we are pleased it has been effective.
Relating back to our response to C23, we think this purpose cannot really be convincingly accomplished
without first stating the lottery model, since the result uses that model. Math is more precise and can be
more convincing than language, the trade off being that some readers won't understand some math. However, the 
math we have in the Intro is so simple we think all readers of Ecology Letters will understand it, so there are
benefits and basically no costs of including it there. } \\
%DONE

\noindent C26) L126: Do you do this in the appendix? \\

\noindent ***\emph{No. We know how it would be done, but we have not formalized the mathematics
in the write-up anywhere. And given length limits and a desire to retain as much simplicity as possible,
we think it's best not to add such mathematics, even to the SI. 
Therefore, to avoid making claims that we have not fully supported
with rigorous mathematics, we rephrased to say the iid assumption is 
``probably not necessary''. Thanks for spurring us to do this.} \\
%DONE

@@

\noindent C27) Eq 2.: I thought we also have to remove the effect of C varying alone and E varying alone? Is this included in the $\#$ term automatically? \\

\noindent ***\emph{No, the equation is correct as is. Please see equation 10 and related equations of 
Ellner et al, 2019, Ecology Letters. Yes, the effects of $C$ and $E$ varying are included in 
the $\#$ term because that terms involves both of them varying (though independently).}

\emph{Also, see our SI section ``Mechanisms other than storage effects" section in SI,
where the following mathematics are presented. 
Since,
\begin{align}
\olep_i^{EC} &= \E[r_i(E_i,C_{i \bs i})]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \olr_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0,
\end{align}
and 
\begin{align}
\olep_i^{(E\shp C)} &= \E[r_i(E_i^\shp,C^\shp_{i \bs i})]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \olrs_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0,
\end{align}
we have
\begin{align}
\olep_i^{(EC)} &= \olep_i^{EC}-\olep_i^{(E\shp C)} \\
&= (\olr_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0) - (\olrs_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0) \\
&= \olr_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0 - \olrs_{i \bs i} + \olep_i^E + \olep_i^C + \epsilon_i^0 \\
&= \olr_{i \bs i} - \olrs_{i \bs i}. 
\end{align}
So,
\begin{align}
\Delta_i^{(EC)} &= \olep_i^{(EC)} - q_{ij}\olep_j^{(EC)} \\
&= (\olr_{i \bs i} - \olrs_{i \bs i}) - q_{ij}(\olr_{j \bs i} - \olrs_{j \bs i}) \\
&= \olr_{i \bs i} - \olrs_{i \bs i} - q_{ij}\olr_{j \bs i} + q_{ij}\olrs_{j \bs i} \\
&= \olr_{i \bs i} - q_{ij}\olr_{j \bs i} - \olrs_{i \bs i} + q_{ij}\olrs_{j \bs i} \\
&= (\olr_{i \bs i} - q_{ij}\olr_{j \bs i}) - (\olrs_{i \bs i} - q_{ij}\olrs_{j \bs i}),
\end{align} 
which is the equation the referee is asking about.} \\
%DONE

@@

\noindent C28) Eq. 6: Maybe highlight that you use different brackets ([EC] versus (EC)), or probably even better, make 5 and 6 one equation using curly under brackets. \\

\noindent ***\emph{Yes, it is a good idea to make it completely clear to the reader that these are 
used distinctly. We now do so, though we do it above this point in the text - it is better, in our view,
to do this the first time we use the hard brackets. So thanks for pointing this out.
We do not know what the referee means by ``using curly under brackets'' so have not done anything
in response to that part of this comment.} \\
%DONE

\noindent C29) L230 $\mu_1-\mu_2$ is negative, why not take $\mu_2-\mu_1$? \\

\noindent ***\emph{Well, we can do this if the editor insists, but to us, avoiding the use of
negative numbers is insufficient reason to have to undertake the substantial and time-consuming 
work, and, more importantly, potential for error, 
of re-doing several figures, and revising many formulas (in the main text and SI).
After all, negative numbers are basic, and represent no impediment to the average ecologist understanding the 
paper. } \\
%DONE

@@

\noindent C30) L316: So why do you show values of $\sigma >6$? And also maybe highlight this in the figure Figure 1d and e: I’m semi convinced by these distributions showing ATA. In 1d it simply appears that the variance scales with the density. 1e is quite unconvincing to me. That’s generally fine, but figure 1 seems to be a conceptual figure, so I wonder whether you don’t have a better example? \\

\noindent ***\emph{Regarding the first referee question listed above, well, what we are 
trying to do here is accurately portray our results about
when ATAs were important and when they were not. To us, the most forthright way of doing that is to
show \textbf{all} the results we generated, including the values of $\sigma$ which are probably unrealistic, and mention
(as we do) that some values are unrealistic. The point of this paragraph is to outline the circumstances
we have explored, using our lottery-model approaches, for which ATAs were important or not, and including all
our results seems important when such is our goal. Furthermore, there's no real cost to showing
all the $\sigma$'s we explored, it does not take up more manuscript space to do so.
Some readers may also take the lesson, for future work, from these explorations that one of the things that probably 
matters for continuing to explore when ATAs are important and when not is probably the strength
of the noise impacting the system. Which seems accurate since Chesson's weak-noise approximations are more likely
to work when noise is weaker. Finally, whether a value of $\sigma$ is ``unrealistic'' is 
a bit of judgement call rather than a formal cutoff, so we hesitate to censor results 
based on such a thing.}

\emph{Regarding the second part of the referee comment, pertaining to Figure 1d and e, it is perplexing to us that 
this is mentioned here,
as the paragraph at L316 is not directly related to Fig. 1d-e. That figure
shows motivational empirical examples, distinct from the lottery model work discussed around L316. Nevertheless, we took
seriously the referee's comment that they were apparently not convinced there are ATAs in Fig. 1d-e.
In fact, one can significance test these results and they are solidly significant - we added
those results, along with the methods used to get them, to a new SI section which is now cited from the
figure caption. We also added some text to the figure caption which we hope will focus the 
readers attention on the visual aspects of these panels which reveal asymmetries of tail association. } \\
%DONE

@@

\noindent C31) Figure 2: I recommend to add a line at $y = 0$. Also the different subgraphs have different $y$-scales which I’m not sure is beneficial, but unsure myself. \\

\noindent ***\emph{There is already a line at $y=0$. You can see it, for instance, on panel (e). 
It gets unavoidably overplotted on panels for which one of the plotted quantities is close to 0. The reason
for different $y$-axis extents is so that you can see the variation within each panel with maximum 
clarity. The main point, here, is to compare the relative contributions of different components
to coexistence, so it's important to show within-plot variation (between lines and across 
values of the $x$-axis) clearly. We tried plotting with the same scales everywhere and it obscured some of the 
patterns. We thank the referee for expressing their own uncertainty, which was
useful to be aware of.} \\
%DONE

\noindent C32) Figure 2 and 3: It appears that ATA is almost never the relevant factor, however I wonder whether this is because ATA is relatively weak, or whether the same analysis of say $\delta_i$ would lead to a similar pattern, it might be worth creating these graphs in the appendix to strengthen your case. \\

\noindent ***\emph{For Fig. 2, the referee is right that ATAs are only rarely important, as explained in the text 
(see first and last paragraphs of section 4.1). However, for Fig 3, the referee is incorrect. In that figure, the ATA term
is usually one of the largest-magnitude contributions to coexistence. Our results as to the importance of ATAs for coexistence 
contrast between our log-normal and beta funcundities lottery models, as explained in the last paragraph
of section 4.1. The contrast is one of the points we are making -- that ATAs are sometimes, but not always, important,
and we need additional work to determine when they are important across a larger collection of systems/models once
those are available.} \\
%DONE

@@

\noindent C33) Figure 2: ATA is always positive, so how can it lead to ATA exclusion? (or ATA rescue in case I misunderstood). \\

\noindent ***\emph{The referee appears to have misinterpreted the figure. The ATA terms is, in fact, usually
negative for this figure, when it is appreciably different from $0$ (the figure used to be Figure 1 in the 
prior submission, but is now figure 3). 
The colors of lines corresponding to different terms are in the legend. And, in the caption,
it is reviewed which of these corresponds to the ATA term. Since these items are clearly labelled, and since 
the $y=0$ line is also already present, we are unsure how to reduce the chances for other readers to make
such errors. But we're happy to hear specific suggestions.} \\
%DONE

@@

\noindent C35) Overall, I agree with the authors that the role of ATAs in coexistence has not been properly considered and the manuscript give convincing results of the importance of ATAs. \\

\noindent ***\emph{Thank you for the positive feedback!} \\
%DONE

\noindent C36) However in order to reach a broad audience of ecologists, the manuscript needs to reach a balance between mathematical definitions and simple explanations. Otherwise, it can remain to technical and of little importance to ecologists interested in the role of environmental variability of community dynamics. I will give some examples where the message can be improved, but I would suggest the authors to do the same effort themselves in other parts of the manuscript. \\

\noindent ***\emph{We have gone to great lengths to address the referee's specific suggestions below, and we have also redoubled our
efforts to make the manuscript as accessible as possible, throughout. We appreciate being spurred to do so, thanks. } \\
%DONE

@@

\noindent C37) At the beginning of the introduction the authors cite previous work to show the prevalence and importance of ATAs for population dynamics, and the stability of ecological communities, and Fig.1 illustrate what ATAs mean in a clear example, yet it is not well explained or at least it was not clear to me how ATA can either promote or hinder coexistence in these initial sentences. \\

\noindent ***\emph{The referee appears to want us to present an intuitive explanation for how ATA effects work, and that 
is certainly a good idea. We have invested substantial efforts into doing so, and think we have succeeded. Though the referee's comment above may 
be interpreted to suggest that they wanted us
to add such an explanation to first paragraph of the paper, the reader is not equiped to understand the explanation at that 
point, so we added a new paragraph later in the Introduction (it is now the
second-to-last paragraph in the Introduction) that begins establishing such an intuition. We also added some text to the end of
the third paragraph of the Introduction that adds to the intuitive understanding of how ATAs can influence coexistence through a concrete 
example (water limitation). We also added a new paragraph to the 
Discussion (the second paragraph of the Discussion). The new Discussion 
paragraph is also supported by two new SI sections and two new SI figures which 
explain how the intuition established in the new main text paragraphs emerges from the modelling (SI sections S11 
and S12, SI Figs S4 and S5).
%DAN: make sure these are the final numbers of these SI sections
We believe that intuition such as the new material establishes should always be based in formal theoretical and 
empirical work, so the focus of the paper remains on our expansion of the MCT and our results showing
the *importance* of ATA effects. Though we definitely think our additions giving an intuition for ATA effects are a large
improvement in the manuscript, so we thank the referee for spurring us to make these additions.  } \\
%DONE except the checking as noted above.

@@

\noindent C38) Then the explanation of the importance of ATAs is done in relationship with the explanation of the storage effect. In detail, the explanation of the storage effect is superb, ... \\

\noindent ***\emph{Thank you for the positive feedback! We have also added some text to the end of that paragraph 
(now the third paragraph of the Introduction) giving a 
concrete case (water limitation) in which one may expect an influence of ATAs. We think the new text definitely
makes it more intuitively reasonable to explore ATA contributions to conexistence.} \\
%DONE

@@

\noindent C39) ...but how ATAs enter in play is not well solved. It is just mentioned that ATAs can modify the EC covariation, which I believe is not enough, a reader might not understand later the explanation when the lottery model is presented. \\

\noindent ***\emph{See our response to C37. We have added substantial material to the Introduction and Discussion that adds an
intuitive explanation for ATAs, and we thank the referee for spurring us to do so. 
We find it unlikely that any focussed reader will have substantial difficulties understanding the lottery model
material in the Introduction, since the lottery model is certainly no more complicated than models such as the Lotka-Volterra 
competition and predator-prey models, which are covered in essentially every introductory undergraduate ecology class, and we can assume
that almost all readers of Ecology Letters have taken such a class. See also comment C25 above, which indicates that another
referee found this material to be compelling.} \\
%DONE

\noindent C40) ATAs can not only modify the importance of the storage effect but also of other mechanisms such as relative non-linearity, which is based on how competition fluctuates in ways that favor the competitor. Could the authors briefly indicate how this could occur? \\

\noindent ***\emph{This cannot occur. ATAs can only influence storage effects, not the other mechanisms, as stated at
the top of section 2.4 and explained mathematically in the SI section cited at the end of section 2.4. Basically, as explained in 2.4, 
this is because ATAs are about covariation between $E$ and $C$, whereas the mechanisms other than storage effects
do not pertain at all to covariation. For instance, relative nonlinearity is about the effects of variation 
in $E$ or $C$, not covariation between them.} \\
%DONE

@@

\noindent C41) The lottery model is used to illustrate the importance of ATAs, and two types of fecundities are taken 
into account (log-normal and beta), can the author explain why both are important, and later on the results and discussion 
why ATAs play a major role with beta fecundities and not with log-normal.
The model used for the chemostat is in its formulation quite different compared to the lottery model. Although the technical
details are left for the supplementary material, can the authors explain more in the detail in the main text how the 
importance of ATAs can be used for this particular model.\\

\noindent ***\emph{This is certainly a good question, and we have gone to substantial lengths to address it. 
Partly in response to this comment and partly in response to earlier comments of referee 3, 
we added substantial new material providing intuitive explanations of ATA effects in the Introduction,
Discussion, and SI (see above responses). That new material also has information comparing the models
and explaining why ATA effects are substantial for the beta-fecundities lottery model and not as strong
for the log-normal fecundities lottery model. These are very positive additions to the manuscript, so we 
thank the referee for spurring us to add them. } \\
%DONE

@@

\noindent C42) The chemostat model is parameterized with information from a previous study, but does this model reproduce observed dynamics of the diatom system? this needs to be shown otherwise the importance of ATAs will be considered for another dynamics that do not correspond to the observed ones. \\

\noindent ***\emph{This was addressed already by the earlier study (Descamps-Julien and Gonzalez, 2005). 
The only parameters we manipulated were the ones pertaining
to the experimentally imposed temperature fluctuations; biological parameters determined, validated and used by Descamps-Julien and Gonzalez
were the same as the biological parameters we used. Descamps-Julien and Gonzalez include results on the extent to
which dynamics predicted by the model were validated by experiment; the validation was not perfect (it could
never be), but was judged to be sufficient, and we agree - see the earlier study for details. 
Outcomes of coexistence/exclusion were 
correctly predicted by the model, among other validations, as described in the earlier paper. 
All we've done is consider, in silico, fluctuations of 
experimentally controlled temperature that sometimes
differ from those actually imposed, in vitro, by the earlier authors. However, the temperature sensitivities of 
biological rates determined and used by the earlier authors (reproduced in SI Fig. S3) span temperatures from 
6-24 degrees C, and all the temperature fluctuations we considered remain within that range. So we
have also not extrapolated beyond the temperature ranges already considered by the earlier paper.} \\
%DONE

@@

\noindent C43) This study novelty is the focus on assymetric tail association between competition and environmental conditions as a mechanism of coexistence. \\

\noindent ***\emph{Thank you for pointing out the novelty.} \\

\noindent C44) The manuscript is very theoretical with a highly technical language hard to follow by a broad range of ecologists. My main suggestion is that the authors need to rewrite it in several parts to make it accessible and convincing. The authors have to make an effort to use a more biological/ecological language with reference to real world potential examples/cases/systems. Some mathematical material could be moved to Supplementary material. \\

\noindent ***\emph{We have now undertaken major additional efforts to make the paper as accessible as possible.
Some of these efforts are summarized in responses to individual referees, above. For instance, we
heavily revised the Introduction to include: a substantial new paragraph that introduces the intuition behind
ATA effects (the second-to-last paragraph of the Intro); and some new material at the end of the third paragraph
of the Intro making it clearer where ATAs can come from using a concrete example based on water limitation.
Efforts to better present the ecological intuition behind ATA effects now also continue in the Discussion, with a new 
paragraph there (now the second paragraph in the Discussion), which is also supported by new some SI sections. }

\emph{We've also made many smaller, but important, changes to the Intro to avoid making readers think they needed
more background knowledge to understand the paper than they actually do. In fact, not much background is
needed. The referees made us aware of, and we have now corrected, a few places early in the Introduction that
made it seem as though readers were expected to know all about storage effects and relative nonlinearity in
competition in order to proceed with the paper. In fact, the readers never needed to know anything about relative 
nonlinearity in competition (that's not needed for the paper at all), and storage effects are explained conceptually
in the Introduction and defined perecisely in Theory, so readers also do not have to have prior knowledge about
storage effects. We apologize for the sentences in the earlier version of the paper which made it seem as though
this was required - we have eliminated those sentences. We have also made explanations of ATAs more accessible,
and eliminated any passages that could be construed as an indication that prior experience with ATAs is 
necessary to understand the paper - it is not.}

\emph{We have also moved to the SI a good deal of content from the Discussion that may have been of interest
more to specialists than generalists,
replacing it with the new paterial helping establish an intuitive sense for how ATAs work (see above).}

\emph{We thank the referees and editor for spurring us to make these changes, as the paper is improved in
its accessibility. The paper is certainly now no less accessible than many other papers I've read recently in 
Ecology Letters, including the two Ellner et al. papers (cited in our study) which helped inspire this 
work and which were published in Ecology Letters.} 
%DONE
